---
title: "FinalProject"
author: "Richard Mailhot"
date: "30 June 2019"
output:
  word_document: default
  html_document: default
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## R Markdown

To create our data to be used, we created the functions mlb_betting_info_clean and hitting_data_clean then combined our data:


```{r}
mlb_betting_info_clean <- function(year){
  
  library('data.table')
  library('schoolmath')
  library('dplyr')
  filename = paste("mlb_odds_20", year, ".csv", sep = "")
  
  mlb_year <- read.csv(filename, stringsAsFactors = FALSE)
  
  ## If we wanted to include the open price
  ##mlb_year$PriceOpen[mlb_year$Open < 0] <- (-100/mlb_year$Open[mlb_year$Open < 0])+1
  ##mlb_year$PriceOpen[mlb_year$Open > 0] <- (mlb_year$Open[mlb_year$Open > 0]/100)+1
  
  mlb_year$PriceClose[mlb_year$Close <0]<- (-100/mlb_year$Close[mlb_year$Close < 0])+1
  mlb_year$PriceClose[mlb_year$Close >0]<- (mlb_year$Close[mlb_year$Close > 0]/100)+1
  
  mlb_year$CloseOdds <- 1/mlb_year$PriceClose
  scores <- c(0:50)


  mlb_year$Win_Even <- mlb_year$Final - shift(mlb_year$Final)
  mlb_year$HandB_Even <-shift(mlb_year$Hand)
  mlb_year$Win_Odd <- mlb_year$Final - shift(mlb_year$Final, type = "lead")
  mlb_year$HandB_Odd <- shift(mlb_year$Hand, type = "lead")
  
  mlb_year$Score_Diff[is.even(mlb_year$Rot)] <- mlb_year$Win_Even[is.even(mlb_year$Rot)]
  mlb_year$Score_Diff[is.odd(mlb_year$Rot)] <- mlb_year$Win_Odd[is.odd(mlb_year$Rot)]
  
  mlb_year$HandB[is.even(mlb_year$Rot)] <- mlb_year$HandB_Even[is.even(mlb_year$Rot)]
  mlb_year$HandB[is.odd(mlb_year$Rot)] <- mlb_year$HandB_Odd[is.odd(mlb_year$Rot)]
  
  mlb_year$Win <- if_else(mlb_year$Score_Diff>0, 1, 0)
  mlb_year$Ones <- 1
  mlb_year$GameNo <- ave(mlb_year$Ones, mlb_year$Team, FUN=cumsum)
  mlb_year <- mlb_year %>% group_by(Date, Team) %>% mutate(DblHead = cumsum(Ones))
  mlb_year$DblHead[mlb_year$DblHead == 1] <- 0
  
  mlb_year$Pitcher <- substring(mlb_year$Pitcher, 2)
  
  mlb_year$Team[mlb_year$Team == 'LOS'] <- 'LAD'
  mlb_year$Date <- as.Date.character(mlb_year$Date, tryFormats = c('%d/%m/%Y', '%Y-%m-%d'))
  mlb_year <- mlb_year[, !(names(mlb_year) %in% c("Win_Odd", "Win_Even", "Close", "Open", "Ones",
                                                  "HandB_Odd", "HandB_Even"))]
  mlb_year
}
```
```{r}

hitting_data_clean <- function(year){
  library('dplyr')
  library('tidyr')
  library('data.table')
  library('schoolmath')
  
  filename = paste("GL20", year, ".csv", sep = "")
  
  hitting_df <- read.csv(filename, stringsAsFactors = FALSE)
  hitting_copy <- hitting_df
  
  hitting_df <- separate(hitting_df, Date, c("Year", "Month_Day"), sep = -4,remove = TRUE,
                           convert = FALSE, extra = "warn", fill = "warn")
  hitting_df <- separate(hitting_df, Month_Day, c("Month", "Day"), sep = -2, remove = TRUE,
                           convert = FALSE, extra = "warn", fill = "warn")
  hitting_df$Date <- as.Date(paste(hitting_df$Day, hitting_df$Month, hitting_df$Year, sep = ".")
                               , format = "%d.%m.%Y")
  hitting_df <- hitting_df[, !(names(hitting_df) %in% c("Month", "Day", "Year"))]
  hitting_df$Home_BB <- hitting_df$Home_BB + hitting_df$Home_HBP
  hitting_df$Visting_BB <- hitting_df$Visting_BB + hitting_df$Visting_HBP
  
  
  hitting_df_visit <- hitting_df[, !(names(hitting_df) %in% c( "Home_RBI",
                                                                     "Home_SacH", "Home_SacFl", "Home_HBP", "Home_IBB", "Home_GDP",
                                                                     "Home_Catcher_Interference", "Home_LOB","Visting_RBI", "Visting_SacH",
                                                                     "Visting_SacFl", "Visting_HBP", "Visting_IBB", "Visting_GDP",
                                                                     "Visting_Catcher_Interference", "Visting_LOB", "Visting_SB",
                                                                     "Visting_CS", "Home_CS", "Home_SB"))]
  hitting_df_home <- hitting_df[, !(names(hitting_df) %in% c("Visting_RBI", "Visting_SacH",
                                                                   "Visting_SacFl", "Visting_HBP", "Visting_IBB", "Visting_GDP",
                                                                   "Visting_Catcher_Interference", "Visting_LOB", "Home_RBI",
                                                                   "Home_SacH", "Home_SacFl", "Home_HBP", "Home_IBB", "Home_GDP",
                                                                   "Home_Catcher_Interference", "Home_LOB", "Visting_SB",
                                                                   "Visting_CS", "Home_CS", "Home_SB"))]
  
  
  column_names_home <- c("DblHead", "DOW", "TeamB", "LeagueB", "GameNoB", "TeamA", "LeagueA", "GameNoA", "Final_ScoreB",
                         "Final_ScoreA", "Length_Outs", "DayNight", "CompletionInfo", "LineB", "LineA", "B_AB", "B_Hits",
                         "B_DB", "B_TR", "B_HR", "B_BB", "B_SO", "A_AB", "A_Hits", "A_DB", "A_TR", "A_HR", "A_BB",
                         "A_SO", "B_SP_ID", "B_SP_Name", "A_SP_ID", "A_SP_Name", "Date")
  
  
  column_names_visit <- c("DblHead", "DOW", "TeamA", "LeagueA", "GameNoA", "TeamB", "LeagueB", "GameNoB", "Final_ScoreA",
                          "Final_ScoreB", "Length_Outs", "DayNight", "CompletionInfo", "LineA", "LineB", "A_AB", "A_Hits",
                          "A_DB", "A_TR", "A_HR", "A_BB", "A_SO", "B_AB", "B_Hits", "B_DB", "B_TR", "B_HR", "B_BB",
                          "B_SO", "A_SP_ID", "A_SP_Name", "B_SP_ID", "B_SP_Name", "Date")
  
  names(hitting_df_home) <- column_names_home
  names(hitting_df_visit) <- column_names_visit
  
  hitting_df_home$VH <- 'H'
  hitting_df_visit$VH <- 'V'
  hitting_df_home$LeagueGameNo <- c(1:nrow(hitting_df_home))
  hitting_df_visit$LeagueGameNo <- c(1:nrow(hitting_df_visit))
  
  
  hitting_df_full <-rbind(hitting_df_home, hitting_df_visit)
  hitting_df_full_copy <- hitting_df_full
  hitting_df_full$A_SP_Name <- substring(hitting_df_full$A_SP_Name, regexpr(" ", hitting_df_full$A_SP_Name) + 1)
  library('stringr')
  hitting_df_full$A_SP_Name <- str_to_upper(hitting_df_full$A_SP_Name)
  
  hitting_df_full$TeamA[hitting_df_full$TeamA == 'ANA'] <- 'LAA'
  hitting_df_full$TeamB[hitting_df_full$TeamB == 'ANA'] <- 'LAA'
  
  hitting_df_full$TeamA[hitting_df_full$TeamA == 'CHA'] <- 'CWS'
  hitting_df_full$TeamB[hitting_df_full$TeamB == 'CHA'] <- 'CWS'
  
  hitting_df_full$TeamA[hitting_df_full$TeamA == 'CHN'] <- 'CUB'
  hitting_df_full$TeamB[hitting_df_full$TeamB == 'CHN'] <- 'CUB'
  
  hitting_df_full$TeamA[hitting_df_full$TeamA == 'KCA'] <- 'KAN'
  hitting_df_full$TeamB[hitting_df_full$TeamB == 'KCA'] <- 'KAN'
  
  hitting_df_full$TeamA[hitting_df_full$TeamA == 'LAN'] <- 'LAD'
  hitting_df_full$TeamB[hitting_df_full$TeamB == 'LAN'] <- 'LAD'
  
  hitting_df_full$TeamA[hitting_df_full$TeamA == 'FLO'] <- 'MIA'
  hitting_df_full$TeamB[hitting_df_full$TeamB == 'FLO'] <- 'MIA'
  
  hitting_df_full$TeamA[hitting_df_full$TeamA == 'NYA'] <- 'NYY'
  hitting_df_full$TeamB[hitting_df_full$TeamB == 'NYA'] <- 'NYY'
  
  hitting_df_full$TeamA[hitting_df_full$TeamA == 'NYN'] <- 'NYM'
  hitting_df_full$TeamB[hitting_df_full$TeamB == 'NYN'] <- 'NYM'
  
  hitting_df_full$TeamA[hitting_df_full$TeamA == 'SDN'] <- 'SDG'
  hitting_df_full$TeamB[hitting_df_full$TeamB == 'SDN'] <- 'SDG'
  
  hitting_df_full$TeamA[hitting_df_full$TeamA == 'SFN'] <- 'SFO'
  hitting_df_full$TeamB[hitting_df_full$TeamB == 'SFN'] <- 'SFO'
  
  hitting_df_full$TeamA[hitting_df_full$TeamA == 'SLN'] <- 'STL'
  hitting_df_full$TeamB[hitting_df_full$TeamB == 'SLN'] <- 'STL'
  
  hitting_df_full$TeamA[hitting_df_full$TeamA == 'TBA'] <- 'TAM'
  hitting_df_full$TeamB[hitting_df_full$TeamB == 'TBA'] <- 'TAM'
  
  hitting_df_full
  
}

```

```{r}
library('dplyr')
library('tidyr')
library('data.table')
library('schoolmath')

source('mlb_betting_info_clean.R')
source('hitting_data_clean.R')

size_df <- data.frame(Year = integer(), Size = integer())

years <- c(15:18)
for (year in years){
  
  
  mlb_betting_df <- mlb_betting_info_clean(year)
  
  hitting_df_full <- hitting_data_clean(year)
  
  
  mlb_df_combined <- merge(mlb_betting_df, hitting_df_full, by.x = c("Team", "Date", "DblHead"), by.y = c("TeamA", "Date", "DblHead"))
  
  mlb_df_combined$difference <- mlb_df_combined$Final-mlb_df_combined$Final_ScoreA
  sum(mlb_df_combined$difference)
  
  mlb_df_combined <- mlb_df_combined[mlb_df_combined$difference == 0, ]
  
  gameNos <- mlb_df_combined$LeagueGameNo[duplicated(mlb_df_combined$LeagueGameNo)]
  
  mlb_df_combined <-mlb_df_combined[ mlb_df_combined$LeagueGameNo %in% gameNos, ]
  
  mlb_df_combined <- mlb_df_combined[order(mlb_df_combined$LeagueGameNo, mlb_df_combined$Rot), ]
  
  mlb_df_combined$cum_AB <- ave(mlb_df_combined$A_AB, mlb_df_combined$Team, FUN = cumsum) - mlb_df_combined$A_AB
  mlb_df_combined$cum_Hits <- ave(mlb_df_combined$A_Hits, mlb_df_combined$Team, FUN = cumsum) - mlb_df_combined$A_Hits
  mlb_df_combined$cum_DB <- ave(mlb_df_combined$A_DB, mlb_df_combined$Team, FUN = cumsum) - mlb_df_combined$A_DB
  mlb_df_combined$cum_TR <- ave(mlb_df_combined$A_TR, mlb_df_combined$Team, FUN = cumsum) - mlb_df_combined$A_TR
  mlb_df_combined$cum_HR <- ave(mlb_df_combined$A_HR, mlb_df_combined$Team, FUN = cumsum) - mlb_df_combined$A_HR
  mlb_df_combined$cum_BB <- ave(mlb_df_combined$A_BB, mlb_df_combined$Team, FUN = cumsum) - mlb_df_combined$A_BB
  mlb_df_combined$cum_SO <- ave(mlb_df_combined$A_SO, mlb_df_combined$Team, FUN = cumsum) - mlb_df_combined$A_SO
  
  mlb_df_combined$A_Ave <- mlb_df_combined$cum_Hits/mlb_df_combined$cum_AB
  mlb_df_combined$A_DB_Ave <- mlb_df_combined$cum_DB/mlb_df_combined$cum_AB
  mlb_df_combined$A_TR_Ave <- mlb_df_combined$cum_TR/mlb_df_combined$cum_AB
  mlb_df_combined$A_HR_Ave <- mlb_df_combined$cum_HR/mlb_df_combined$cum_AB
  mlb_df_combined$A_BB_Ave <- mlb_df_combined$cum_BB/mlb_df_combined$cum_AB
  mlb_df_combined$A_SO_Ave <- mlb_df_combined$cum_SO/mlb_df_combined$cum_AB
  
  mlb_df_combined$cum_AB <- ave(mlb_df_combined$B_AB, mlb_df_combined$TeamB, FUN = cumsum) - mlb_df_combined$B_AB
  mlb_df_combined$cum_Hits <- ave(mlb_df_combined$B_Hits, mlb_df_combined$TeamB, FUN = cumsum) - mlb_df_combined$B_Hits
  mlb_df_combined$cum_DB <- ave(mlb_df_combined$B_DB, mlb_df_combined$TeamB, FUN = cumsum) - mlb_df_combined$B_DB
  mlb_df_combined$cum_TR <- ave(mlb_df_combined$B_TR, mlb_df_combined$TeamB, FUN = cumsum) - mlb_df_combined$B_TR
  mlb_df_combined$cum_HR <- ave(mlb_df_combined$B_HR, mlb_df_combined$TeamB, FUN = cumsum)- mlb_df_combined$B_HR
  mlb_df_combined$cum_BB <- ave(mlb_df_combined$B_BB, mlb_df_combined$TeamB, FUN = cumsum)- mlb_df_combined$B_BB
  mlb_df_combined$cum_SO <- ave(mlb_df_combined$B_SO, mlb_df_combined$TeamB, FUN = cumsum)- mlb_df_combined$B_SO
  
  mlb_df_combined$B_Ave <- mlb_df_combined$cum_Hits/mlb_df_combined$cum_AB
  mlb_df_combined$B_DB_Ave <- mlb_df_combined$cum_DB/mlb_df_combined$cum_AB
  mlb_df_combined$B_TR_Ave <- mlb_df_combined$cum_TR/mlb_df_combined$cum_AB
  mlb_df_combined$B_HR_Ave <- mlb_df_combined$cum_HR/mlb_df_combined$cum_AB
  mlb_df_combined$B_BB_Ave <- mlb_df_combined$cum_BB/mlb_df_combined$cum_AB
  mlb_df_combined$B_SO_Ave <- mlb_df_combined$cum_SO/mlb_df_combined$cum_AB
  
  mlb_df_combined$cum_AB <- ave(mlb_df_combined$B_AB, mlb_df_combined$A_SP_ID, FUN = cumsum)- mlb_df_combined$B_AB
  mlb_df_combined$cum_Hits <- ave(mlb_df_combined$B_Hits, mlb_df_combined$A_SP_ID, FUN = cumsum)- mlb_df_combined$B_Hits
  mlb_df_combined$cum_DB <- ave(mlb_df_combined$B_DB, mlb_df_combined$A_SP_ID, FUN = cumsum) - mlb_df_combined$B_DB
  mlb_df_combined$cum_TR <- ave(mlb_df_combined$B_TR, mlb_df_combined$A_SP_ID, FUN = cumsum) - mlb_df_combined$B_TR
  mlb_df_combined$cum_HR <- ave(mlb_df_combined$B_HR, mlb_df_combined$A_SP_ID, FUN = cumsum) - mlb_df_combined$B_HR
  mlb_df_combined$cum_BB <- ave(mlb_df_combined$B_BB, mlb_df_combined$A_SP_ID, FUN = cumsum) - mlb_df_combined$B_BB
  mlb_df_combined$cum_SO <- ave(mlb_df_combined$B_SO, mlb_df_combined$A_SP_ID, FUN = cumsum) - mlb_df_combined$B_SO
  
  mlb_df_combined$P_A_Ave <- mlb_df_combined$cum_Hits/mlb_df_combined$cum_AB
  mlb_df_combined$P_A_DB_Ave <- mlb_df_combined$cum_DB/mlb_df_combined$cum_AB
  mlb_df_combined$P_A_TR_Ave <- mlb_df_combined$cum_TR/mlb_df_combined$cum_AB
  mlb_df_combined$P_A_HR_Ave <- mlb_df_combined$cum_HR/mlb_df_combined$cum_AB
  mlb_df_combined$P_A_BB_Ave <- mlb_df_combined$cum_BB/mlb_df_combined$cum_AB
  mlb_df_combined$P_A_SO_Ave <- mlb_df_combined$cum_SO/mlb_df_combined$cum_AB
  
  
  mlb_df_combined$cum_AB <- ave(mlb_df_combined$A_AB, mlb_df_combined$B_SP_ID, FUN = cumsum) - mlb_df_combined$A_AB
  mlb_df_combined$cum_Hits <- ave(mlb_df_combined$A_Hits, mlb_df_combined$B_SP_ID, FUN = cumsum) - mlb_df_combined$A_Hits
  mlb_df_combined$cum_DB <- ave(mlb_df_combined$A_DB, mlb_df_combined$B_SP_ID, FUN = cumsum) - mlb_df_combined$A_DB
  mlb_df_combined$cum_TR <- ave(mlb_df_combined$A_TR, mlb_df_combined$B_SP_ID, FUN = cumsum) - mlb_df_combined$A_TR
  mlb_df_combined$cum_HR <- ave(mlb_df_combined$A_HR, mlb_df_combined$B_SP_ID, FUN = cumsum) - mlb_df_combined$A_HR
  mlb_df_combined$cum_BB <- ave(mlb_df_combined$A_BB, mlb_df_combined$B_SP_ID, FUN = cumsum) - mlb_df_combined$A_BB
  mlb_df_combined$cum_SO <- ave(mlb_df_combined$A_SO, mlb_df_combined$B_SP_ID, FUN = cumsum) - mlb_df_combined$A_SO
  
  mlb_df_combined$P_B_Ave <- mlb_df_combined$cum_Hits/mlb_df_combined$cum_AB
  mlb_df_combined$P_B_DB_Ave <- mlb_df_combined$cum_DB/mlb_df_combined$cum_AB
  mlb_df_combined$P_B_TR_Ave <- mlb_df_combined$cum_TR/mlb_df_combined$cum_AB
  mlb_df_combined$P_B_HR_Ave <- mlb_df_combined$cum_HR/mlb_df_combined$cum_AB
  mlb_df_combined$P_B_BB_Ave <- mlb_df_combined$cum_BB/mlb_df_combined$cum_AB
  mlb_df_combined$P_B_SO_Ave <- mlb_df_combined$cum_SO/mlb_df_combined$cum_AB
  
  mlb_df_combined$P_D_Ave <- mlb_df_combined$P_A_Ave - mlb_df_combined$P_B_Ave
  mlb_df_combined$P_D_DB_Ave <- mlb_df_combined$P_A_DB_Ave - mlb_df_combined$P_B_DB_Ave
  mlb_df_combined$P_D_TR_Ave <- mlb_df_combined$P_A_TR_Ave - mlb_df_combined$P_B_TR_Ave
  mlb_df_combined$P_D_HR_Ave <- mlb_df_combined$P_A_HR_Ave - mlb_df_combined$P_B_HR_Ave
  mlb_df_combined$P_D_BB_Ave <- mlb_df_combined$P_A_BB_Ave - mlb_df_combined$P_B_BB_Ave
  mlb_df_combined$P_D_SO_Ave <- mlb_df_combined$P_A_SO_Ave - mlb_df_combined$P_B_SO_Ave
  
  mlb_df_combined$D_Ave <- mlb_df_combined$A_Ave - mlb_df_combined$B_Ave
  mlb_df_combined$D_DB_Ave <- mlb_df_combined$A_DB_Ave - mlb_df_combined$B_DB_Ave
  mlb_df_combined$D_TR_Ave <- mlb_df_combined$A_TR_Ave - mlb_df_combined$B_TR_Ave
  mlb_df_combined$D_HR_Ave <- mlb_df_combined$A_HR_Ave - mlb_df_combined$B_HR_Ave
  mlb_df_combined$D_BB_Ave <- mlb_df_combined$A_BB_Ave - mlb_df_combined$B_BB_Ave
  mlb_df_combined$D_SO_Ave <- mlb_df_combined$A_SO_Ave - mlb_df_combined$B_SO_Ave
  
  columns <- c(1, 2, 4, 5, 6, 7, 8, 9, 10, 11, 12, 13, 14, 16, 45, 80:91)
  
  
  mlb_df_reg <- mlb_df_combined[,columns]
  mlb_df_reg$Home[mlb_df_reg$VH.x == 'H'] <- 1
  mlb_df_reg$Home[mlb_df_reg$VH.x == 'V'] <- 0
  
  size_this_year <- data.frame(Year = year, Size = nrow(mlb_df_reg))
  size_df <- rbind(size_df, size_this_year)
  
  write.csv(mlb_df_reg, file = paste("mlb_data_combined_20", year, ".csv", sep = ""), row.names = FALSE)
  
}

```

We also ran the above slightly modified to produce offensive averages for the throwing hand of the pitcher a team is facing.

Once these csvs were created we built our regression. The below code was ran with 2015, 2016 and 2017 as the training data and 2016, 2017 and 2018 as the test data. Only one is shown for the sake of space.

```{r}
mlb_train <- read.csv('mlb_data_combined_wPH_2015.csv')
mlb_test <- read.csv('mlb_data_combined_wPH_2016.csv')

mlb_train <- mlb_train[complete.cases(mlb_train), ]
mlb_test <- mlb_test[complete.cases(mlb_test), ]
#mlb_2017_train$Loss[mlb_2017_train$Win == 0] <-1
#mlb_2017_train$Loss[mlb_2017_train$Win == 1] <-0

fit <- glm(Win ~ Home + D_Ave + D_DB_Ave + D_TR_Ave + D_HR_Ave + D_BB_Ave + D_SO_Ave + P_D_Ave + P_D_DB_Ave + P_D_TR_Ave + P_D_HR_Ave + P_D_BB_Ave + P_D_SO_Ave , data = mlb_train)
summary(fit)
step(fit)

# Gives Win ~ Home + D_DB_Ave + D_TR_Ave + D_BB_Ave + P_D_BB_Ave + P_D_SO_Ave for 2017

# Gives Win ~ Home + D_DB_Ave + D_TR_Ave + D_BB_Ave + D_SO_Ave + P_D_Ave + P_D_DB_Ave 
# + P_D_HR_Ave + P_D_BB_Ave for 2016

# Gives Win ~ Home + D_DB_Ave + D_BB_Ave + P_D_TR_Ave + P_D_BB_Ave for 2015

fit2 <- glm(formula = Win ~ Home + D_DB_Ave + D_TR_Ave + D_BB_Ave + P_D_TR_Ave + 
              P_D_SO_Ave, data = mlb_train)
summary(fit2)
mlb_test$reg_predict <- predict(fit2, mlb_test, type = "response")
mlb_test$reg_Bet <- ((mlb_test$reg_predict*(mlb_test$PriceClose-1)-(1-mlb_test$reg_predict))/(mlb_test$PriceClose-1))*1000
mlb_test$reg_Bet[mlb_test$CloseOdds>mlb_test$reg_predict] <- 0
mlb_test$reg_Winnings <- (mlb_test$Win*mlb_test$PriceClose - 1)*mlb_test$reg_Bet


sum(mlb_test$reg_Winnings)

# mlb_test$WinningsEven[mlb_test$CloseOdds<mlb_test$predict] <- 100*mlb_test $Win[mlb_test$CloseOdds<mlb_test$predict]*mlb_test$PriceClose[mlb_test$CloseOdds<mlb_test$predict]-100
# mlb_test$WinningsEven[mlb_test$CloseOdds>mlb_test$predict] <- 0
# 
# sum(mlb_test$WinningsEven)
mlb_reg_summary <- mlb_test[, c(1:3, 7:10, 12, 29:31)]

write.csv(mlb_reg_summary, file = 'regression_betting_performance_2016.csv', row.names = FALSE)

```


